Comparing Hydrogen and Jet-A for an N+3 Turbofan with Water Recirculation using Gradient-Based Optimization

### 2023-05-10 conference-paper/main.tex
addressed changes and updated slides

% Message 2: Background and references to support using H2 and water injection in HBTF engines
% [x] TODO: AL-PA, Need to find a journal/conference paper to cite for the following statement.  Maybe find one or more papers other than just the NASA report?
Water injection is the process of introducing water upstream of the combustor as finely atomized droplets.
NASA, Boeing, and Rolls-Royce studied this concept and suggested that this technique reduces the NOx emissions as much as 47 percent~\cite{Daggett2010}.
Additionally, water injection improves fuel efficiency and thrust output with lower combustion temperatures that can improve the lifetime of turbine blades and reduce noise~\cite{Daggett2010}.
% [x] TODO: AL-PA, Should try to find a citation for this if one exists.
Traditional propulsion systems that burn hydrocarbon fuels would require external demineralized water storage on the aircraft for water injection~\cite{Mourouzidis2015}.
The added weight of tanks, pumping, and ducting makes this concept infeasible for a conventional aircraft.
The main product of hydrogen combustion is water vapor.
We can recover the vapor from the exhaust stream~\cite{Strom2002} with condensers and recirculate it upstream to eliminate additional storage requirements.
% Condensing water vapor from the exhaust stream of hydrogen combustion and recirculating it eliminates any additional storage requirements.
This system represents a closed loop water feedback system inside the propulsion cycle.
% This allows for the theoretical design of a closed loop water feedback system inside the propulsion cycle.
Pratt and Whitney are actively researching this technology to increase the efficiency of hydrogen-powered propulsion~\cite{arpa-e_2021}.
% They suggest that water vapor can be recovered by utilizing the thermal properties of hydrogen to condense some of the water in the turbofan exhaust.

@@

% [x] TODO: AL-PA, Remove the black box with the component descriptions from the bottom left of the figure.
% [x] TODO: AL-PA, Add the description of the figure elements to the caption.
% [x] TODO: AL-PA, Consider changing the caption to "Simplified layout of the N+3 engine cycle, adapted from~\citet{Hendricks2019}"
% [x] TODO: AL-PA, Remove the flight conditions component and add the bypass duct component below the splitter. I think the bypass is an important enough aspect of the engine to include.
% [ ] TODO: AL-PA, Missing station labels for the bypass duct, splitter, and inlet
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=1.0\textwidth]{N3_cycle.pdf}
  \caption{
    Simplified layout of the N+3 engine cycle, adapted from~\citet{Hendricks2019}.
    Black arrows are flow connections, red lines are mechanical connections, blue boxes are cycle elements, red boxes are shaft elements.
    Flow stations are identified with bubble text.
  }
  \label{fig:N3_original}
\end{figure}

@@

% [x] TODO: AL-PA, simplify this XDSM to show just the optimization setup. Condense the MDA into a single block and remove all the multipoint acronyms.
% [ ] TODO: AL-PA, Combine the TOC and Off design stack into a single block that just says (MDP Cycle Model)
% [ ] TODO: AL-PA, Remove the map scalar, core area, and water recovery area connections
% [ ] TODO: AL-PA, The symbol for thrust is "F" not "T"
% [ ] TODO: AL-PA, Show the final outputs on the left side of the XDSM and include a superscript "*" to denote optimal values. i.e \dot{m}_\text{fuel}^*
\begin{figure}[!hbt]
  \centering
  \includegraphics[width=0.5\textwidth]{N3_opt_XDSM.pdf}
  \caption{
    XDSM diagram of the multipoint optimization problem with objective, constraints, and MDA block.
    The XDSM diagram shows the variables and outputs within the optimization model and how each of these values is connected to the optimizer and MDA block.}
  \label{fig:N3_opt_xdsm}
\end{figure}

@@

\begin{table}[hbt!]
  \centering
  \caption{Optimization results of the N+3 engine for four optimizations using Jet-A and hydrogen fuel with and without water recirculation.
  }
  \small
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{r l l l l l l}
                & Variable/Function         & Jet-A Dry & Jet-A Wet & \ce{H2} Dry & \ce{H2} Wet & Units           \\
    \toprule
    objective   & $\dot{m}_{\rm{fuel,CRZ}}$ & 0.639     & 0.606     & 0.233       & 0.221       & \si{lbm/s}      \\
                & $\rm{TSEC}_{\rm{CRZ}}$    & 8186.254  & 7766.338  & 8295.236    & 7859.635    & \si{lbm/hr/lbf} \\
    \hline
    variables   & $x_{H2O,CRZ}$             & -         & 0.3       & -           & 0.17        & -               \\
                & $\rm{PR}_{fan,TOC}$       & 1.285     & 1.292     & 1.288       & 1.297       & -               \\
                & $\rm{PR}_{LPC,TOC}$       & 4.0       & 4.0       & 4.0         & 4.0         & -               \\
                & $\rm{OPR}_{\rm{TOC}}$     & 61.012    & 61.188    & 60.660      & 61.952      & -               \\
                & $T_{4,ratio}$             & 0.913     & 0.911     & 0.918       & 0.915       & -               \\
                & $T_{4,RTO}$               & 3220.574  & -         & 3204.071    & -           & $^\circ$R       \\
                & $V_{ratio,CRZ}$           & 1.35      & 1.35      & 1.35        & 1.35        & -               \\
    \hline
    constraints & $F_{net,TOC}$             & 5800.0    & 5800.0    & 5800.0      & 5800.0      & lbf             \\
                & $D_{Fan}$                 & 100.0     & 100.0     & 100.0       & 100.0       & $in^2$          \\
    \bottomrule
  \end{tabular}
  \label{tab:dv_opt}
\end{table}

@@

With the addition of water recovery, the propulsive efficiency increases and leads to higher fan pressure ratios and lower TSEC for both fuels.
Additionally, the $T_{4,\text{ratio}}$ is reduced with water recovery leading to cooler combustion temperatures at TOC and subsequently lower fuel burn at CRZ.
We observe the optimal $T_{4,\text{RTO}}$ is larger for Jet-A than hydrogen because hydrogen has a larger combustion flammability limit and the engine can burn at a lower equivalence ratio to attain the same thrust requirements~\cite{Adler2023}.
% [ ] TODO: AL-PA, why is this favorable for emissions reduction?  Need to cite or explain.
Burning leaner reduces the fuel consumption and is favorable for emissions reduction.
% The additional water in the combustion improves the propulsive efficiency and enables the increase in overall pressure ratio.
% As noted earlier, the water recovery fractions are different because the optimizer pushes $X_{\rm{{H_2O},CRZ}}$ to the highest value it can attain.
% The $\rm{PR}_{fan,TOC}$ and $V_{ratio,CRZ}$ design variables go to their upper and lower limits, respectively.
$V_\text{ratio,CRZ}$ is at the lower bound for all four cases to maximize the amount of mass flow through the bypass stream.

% NOTE: PA-PA/AL replace this with a parallel axis plot?
% [x] TODO: AL-PA, Yeah let's do the parallel axis plot for the basic thermodynamic parameters
% AL-PA, We can make another plot for the water fractions and the dpQp study.
% [ ] TODO: AL-PA, Can you make the y-axis range for T4_ratio smaller so we can see the differences a bit better.  Maybe like 0.85 - 0.95
% [ ] TODO: AL-PA, Same comment as above but for FPR, maybe 1.28-1.32
% [ ] TODO: AL-PA, If a variable has bounds, can we make the y-axis limits match the bounds?
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=1.0\textwidth]{N3_parallel_coords.pdf}
  \caption{Optimization results of the N+3 engine with and without no water recovery using Jet-A and H2 as the fuel.
    The red lines show Jet-A fuel and the blue lines show \ce{H2} fuel.
    The solid lines show the engine without water recovery and the dashed lines show the engine with water recovery.}
  \label{fig:parallel_coords}
\end{figure}

@@

% [x] TODO: I would change this to the horizontal bar chart and plot both H2 and Jet-A on the same axis
% See this example in niceplots: https://github.com/mdolab/niceplots/blob/main/examples/plot_bar_chart.py and https://github.com/mdolab/niceplots/blob/main/examples/ref/bar_chart.png
% [ ] TODO: Can you add a line to bar chart that shows the improvement between baseline Jet-A and hydrogen with water recovery?
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=0.65\textwidth]{bar_chart.pdf}
  \caption{Percent relative improvement in thrust specific energy consumption (TSEC) of the N+3 engine with water recovery.
    This plot compares the relative improvement of the two optimization problems with water recovery compared to the optimization problems without water recovery.}
  \label{fig:barchart}
\end{figure}

The TSEC improvement at CRZ due to water recovery for both fuels is shown on a bar chart in Figure~\ref{fig:barchart}.
Water recovery improves the TSEC approximately 5\% for both Jet-A and hydrogen fueled cycles.
% We see an improvement greater than 5\% at CRZ for TSEC for both fuels.
% [x] TODO: AL-PA, This is a good point, but maybe combine with the paragraph after the figure?
Most of the design variables for the optimal designs with and without water recovery vary only slightly for both fuels.
This suggests that the water recovery system accounts for a majority of the efficiency improvements.
Achieving a 5\% reduction in TSEC represents a large fuel savings over the lifecycle of an engine.
% This level of improvement would have a significant reduction in fuel costs over the course of an engine's lifecycle.
% [x] TODO: AL-PA, Add a sentence that introduces this citation and paraphrases the claim.  Then relate it to your findings.
As~\citeauthor{Strom2002} point out, a hydrogen-fueled engine would have close to 3 times more water in the exhaust than a similar Jet-A fueled counterpart.
% [ ] TODO: AL-PA, I think we need to add some numbers here for the mass flow rate of water for jet-a and hydrogen to help this claim.
% Therefore, \ce{H2} is a much more appealing fuel to use with the proposed closed-loop water recovery system.
Hydrogen combustion offers an immediate reduction in emissions and produces more water in the exhaust stream.
With an advanced water recovery system, we can achieve better TSEC with hydrogen fuel than Jet-A.
To make hydrogen fueled engines a reality, we need to take advantage of the additional benefits of hydrogen combustion.
\subsection{Condenser Design Space Study}
\label{sub:dpqp_sweep}
% [x] TODO: AL-PA, I would emphasize that we did this sweep to explore the design space following our optimization results.
%                  Might be worth moving this into a new subsection.
% [x] TODO: AL-PA, This study wasn't motivated by a lack of literature, we want to explore the potential design space for the technology we just proposed and analyzed.
The efficiency improvement gained from water recirculation warrant a design space study for potential condenser technology.
Companies that are developing engines with water recirculation suggest that hydrogen can be used to as the heat sink for a water condenser~\cite{arpa-e_2021}.
We consider a condenser designed to fit inside the exhaust stream of a hydrogen-fueled turbofan engine.
The trade space for a condenser in the exhaust stream depends on the pressure loss and water recovery capability.
% A low pressure loss condenser would need to be developed for the water recirculation loop to be implemented.
% [x] TODO: AL-PA, dPqP is the relative total pressure loss not just a "pressure loss"

@@

% [ ] TODO: AL-PA, Re-write this part to match the new figure that shows the mass flow of water extracted (per our slack convo)
The total water available in the core stream as a function of the same pressure loss is shown in Figure~\ref{fig:dpqp_wdot}.
Figure~\ref{fig:dpqp_wdot} shows that the optimizer finds designs that increase the available water in the core stream.
Since the water recovery fraction is always pushed to 17\%, the water mass flow rate that is recovered increases to overcome the pressure loss.

@@

% NOTE: PA-AL, any other plots to include?
% AL, Nah we are good to go with plots

@@

We performed a design space study to understand the technology requirements for a condenser that might operate in the proposed water recovery system.
% There are technological obstacles to overcome in order to implement this water recirculation system in a modern turbofan engine.
% The main obstacle would be how to design a condenser that can recover a portion of the exhaust water without a excessive drag pressure losses.
% Therefore, a condenser design space study was completed to show the feasible design space for such a condenser.
The results showed that a hydrogen-fueled cycle with water recovery can sustain a pressure loss of 16\% across a condenser before the efficiency benefits diminish.
% The results of the study showed that at the maximum possible water recovery fraction for hydrogen quite a large pressure loss can be imposed before the improvements are canceled out.
% Furthermore, using hydrogen as the fuel allows for designs that are not available for an engine using Jet-A.
% [ ] TODO: AL-PA, I suggest you bolster the discussion of these two sentences in the results section and maybe add a citation?
The thermal properties of hydrogen provide a heat sink and thus a readily available resource for water condensation within the condenser.
The hydrogen fuel absorbs heat and is then used in the combustion process, further improving engine efficiency.
% The novel water recovery system shows promising efficiency results when used with hydrogen fuel.
Hydrogen-fueled UHBR turbofan engines can be more efficient than Jet-A counterparts when optimized considering water recovery.
Taking advantage of the beneficial properties of hydrogen is paramount for new engine technologies that can reduce emissions while improving performance.
% This water recovery system provides a potential avenue for hydrogen engine design that could help overcome the technological issues associated with hydrogen as a fuel.

>>>>

% Message 2: Background and references to support using H2 and water injection in HBTF engines
% [x] TODO: AL-PA, Need to find a journal/conference paper to cite for the following statement.  Maybe find one or more papers other than just the NASA report?
Water injection is the process of introducing water upstream of the combustor as finely atomized droplets.
NASA, Boeing, and Rolls-Royce studied this concept and suggested that this technique reduces the NOx emissions as much as 47 percent~\cite{Daggett2010}.
Additionally, water injection improves fuel efficiency and thrust output with lower combustion temperatures that can improve the lifetime of turbine blades and reduce noise~\cite{Daggett2010}.
% [x] TODO: AL-PA, Should try to find a citation for this if one exists.
Traditional propulsion systems that burn hydrocarbon fuels would require external demineralized water storage on the aircraft for water injection~\cite{Mourouzidis2015}.
The added weight of tanks, pumping, and ducting makes this concept infeasible for a conventional aircraft.
The main product of hydrogen combustion is water vapor~\cite{Strom2002}.
We can recover the vapor from the exhaust stream with condensers and recirculate it upstream to eliminate additional storage requirements.
If hydrogen is used as the fuel then it can be used as a heat sink for the condenser~\cite{Boggia2002}.
% Condensing water vapor from the exhaust stream of hydrogen combustion and recirculating it eliminates any additional storage requirements.
This system represents a closed loop water feedback system inside the propulsion cycle.
% This allows for the theoretical design of a closed loop water feedback system inside the propulsion cycle.
Pratt and Whitney are actively researching this technology to increase the efficiency of hydrogen-powered propulsion~\cite{arpa-e_2021}.
% They suggest that water vapor can be recovered by utilizing the thermal properties of hydrogen to condense some of the water in the turbofan exhaust.

@@

% [x] TODO: AL-PA, Remove the black box with the component descriptions from the bottom left of the figure.
% [x] TODO: AL-PA, Add the description of the figure elements to the caption.
% [x] TODO: AL-PA, Consider changing the caption to "Simplified layout of the N+3 engine cycle, adapted from~\citet{Hendricks2019}"
% [x] TODO: AL-PA, Remove the flight conditions component and add the bypass duct component below the splitter. I think the bypass is an important enough aspect of the engine to include.
% [x] TODO: AL-PA, Missing station labels for the bypass duct, splitter, and inlet
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=1.0\textwidth]{N3_cycle.pdf}
  \caption{
    Simplified layout of the N+3 engine cycle, adapted from~\citet{Hendricks2019}.
    Black arrows are flow connections, red lines are mechanical connections, blue boxes are cycle elements, red boxes are shaft elements.
    Flow stations are identified with bubble text.
  }
  \label{fig:N3_original}
\end{figure}

@@

% [x] TODO: AL-PA, simplify this XDSM to show just the optimization setup. Condense the MDA into a single block and remove all the multipoint acronyms.
% [x] TODO: AL-PA, Combine the TOC and Off design stack into a single block that just says (MDP Cycle Model)
% [x] TODO: AL-PA, Remove the map scalar, core area, and water recovery area connections
% [x] TODO: AL-PA, The symbol for thrust is "F" not "T"
% [x] TODO: AL-PA, Show the final outputs on the left side of the XDSM and include a superscript "*" to denote optimal values. i.e \dot{m}_\text{fuel}^*
\begin{figure}[!hbt]
  \centering
  \includegraphics[width=0.5\textwidth]{N3_opt_XDSM.pdf}
  \caption{
    XDSM diagram of the multipoint optimization problem with objective, constraints, and MDA block.
    The XDSM diagram shows the variables and outputs within the optimization model and how each of these values is connected to the optimizer and MDA block.}
  \label{fig:N3_opt_xdsm}
\end{figure}

@@

\begin{table}[hbt!]
  \centering
  \caption{Optimization results of the N+3 engine for four optimizations using Jet-A and hydrogen fuel with and without water recirculation.
  }
  \small
  \renewcommand{\arraystretch}{1.2}
  \begin{tabular}{r l l l l l l}
                & Variable/Function         & Jet-A Dry & Jet-A Wet & \ce{H2} Dry & \ce{H2} Wet & Units      \\
    \toprule
    objective   & $\dot{m}_{\rm{fuel,CRZ}}$ & 0.639     & 0.606     & 0.233       & 0.221       & \si{lbm/s} \\
    % & $\rm{TSEC}_{\rm{CRZ}}$    & 8186.254  & 7766.338  & 8295.236    & 7859.635    & \si{lbm/hr/lbf} \\
    \hline
    variables   & $x_{H2O,CRZ}$             & -         & 0.3       & -           & 0.17        & -          \\
                & $\rm{PR}_{fan,TOC}$       & 1.285     & 1.292     & 1.288       & 1.297       & -          \\
                & $\rm{PR}_{LPC,TOC}$       & 4.0       & 4.0       & 4.0         & 4.0         & -          \\
                & $\rm{OPR}_{\rm{TOC}}$     & 61.012    & 61.188    & 60.660      & 61.952      & -          \\
                & $T_{4,ratio}$             & 0.913     & 0.911     & 0.918       & 0.915       & -          \\
                & $T_{4,RTO}$               & 3220.574  & -         & 3204.071    & -           & $^\circ$R  \\
                & $V_{ratio,CRZ}$           & 1.35      & 1.35      & 1.35        & 1.35        & -          \\
    \hline
    constraints & $F_{net,TOC}$             & 5800.0    & 5800.0    & 5800.0      & 5800.0      & lbf        \\
                & $D_{Fan}$                 & 100.0     & 100.0     & 100.0       & 100.0       & $in^2$     \\
    \bottomrule
  \end{tabular}
  \label{tab:dv_opt}
\end{table}

@@

With the addition of water recovery, the propulsive efficiency increases and leads to higher fan pressure ratios and lower TSEC for both fuels.
Additionally, the $T_{4,\text{ratio}}$ is reduced with water recovery leading to cooler combustion temperatures at TOC and subsequently lower fuel burn at CRZ.
We observe the optimal $T_{4,\text{RTO}}$ is larger for Jet-A than hydrogen because hydrogen has a larger combustion flammability limit and the engine can burn at a lower equivalence ratio to attain the same thrust requirements~\cite{Adler2023}.
% [x] TODO: AL-PA, why is this favorable for emissions reduction?  Need to cite or explain.
Burning leaner lowers the FAR which reduces the fuel consumption for a given engine flow rate and is favorable for emissions reduction.
% The additional water in the combustion improves the propulsive efficiency and enables the increase in overall pressure ratio.
% As noted earlier, the water recovery fractions are different because the optimizer pushes $X_{\rm{{H_2O},CRZ}}$ to the highest value it can attain.
% The $\rm{PR}_{fan,TOC}$ and $V_{ratio,CRZ}$ design variables go to their upper and lower limits, respectively.
$V_\text{ratio,CRZ}$ is at the lower bound for all four cases to maximize the amount of mass flow through the bypass stream.

% NOTE: PA-PA/AL replace this with a parallel axis plot?
% [x] TODO: AL-PA, Yeah let's do the parallel axis plot for the basic thermodynamic parameters
% AL-PA, We can make another plot for the water fractions and the dpQp study.
% [x] TODO: AL-PA, Can you make the y-axis range for T4_ratio smaller so we can see the differences a bit better.  Maybe like 0.85 - 0.95
% [x] TODO: AL-PA, Same comment as above but for FPR, maybe 1.28-1.32
% [x] TODO: AL-PA, If a variable has bounds, can we make the y-axis limits match the bounds?
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=1.0\textwidth]{N3_parallel_coords.pdf}
  \caption{Optimization results of the N+3 engine with and without no water recovery using Jet-A and H2 as the fuel.
    The red lines show Jet-A fuel and the blue lines show \ce{H2} fuel.
    The solid lines show the engine without water recovery and the dashed lines show the engine with water recovery.}
  \label{fig:parallel_coords}
\end{figure}

@@

% [x] TODO: I would change this to the horizontal bar chart and plot both H2 and Jet-A on the same axis
% See this example in niceplots: https://github.com/mdolab/niceplots/blob/main/examples/plot_bar_chart.py and https://github.com/mdolab/niceplots/blob/main/examples/ref/bar_chart.png
% [x] TODO: Can you add a line to bar chart that shows the improvement between baseline Jet-A and hydrogen with water recovery?
\begin{figure}[hbt!]
  \centering
  \includegraphics[width=0.65\textwidth]{bar_chart.pdf}
  \caption{Percent relative improvement in thrust specific energy consumption (TSEC) of the N+3 engine with water recovery.
    This plot compares the relative improvement of the two optimization problems with water recovery compared to the optimization problems without water recovery.}
  \label{fig:barchart}
\end{figure}

The TSEC improvement at CRZ due to water recovery for both fuels is shown on a bar chart in Figure~\ref{fig:barchart}.
Water recovery improves the TSEC approximately 5\% for both Jet-A and hydrogen fueled cycles.
% We see an improvement greater than 5\% at CRZ for TSEC for both fuels.
% [x] TODO: AL-PA, This is a good point, but maybe combine with the paragraph after the figure?
Most of the design variables for the optimal designs with and without water recovery vary only slightly for both fuels.
This suggests that the water recovery system accounts for a majority of the efficiency improvements.
Achieving a 5\% reduction in TSEC represents a large fuel savings over the lifecycle of an engine.
% This level of improvement would have a significant reduction in fuel costs over the course of an engine's lifecycle.
% [x] TODO: AL-PA, Add a sentence that introduces this citation and paraphrases the claim.  Then relate it to your findings.
As~\citeauthor{Strom2002} point out, a hydrogen-fueled engine would have close to 3 times more water in the exhaust than a similar Jet-A fueled counterpart.
% [x] TODO: AL-PA, I think we need to add some numbers here for the mass flow rate of water for jet-a and hydrogen to help this claim.
The engine designs with water recovery showed that there was about 2.38\si{lbm/s} and 1.08\si{lbm/s} of water in the core stream for \ce{H2} and Jet-A, respectively
% Therefore, \ce{H2} is a much more appealing fuel to use with the proposed closed-loop water recovery system.
Hydrogen combustion offers an immediate reduction in emissions and produces more water in the exhaust stream.
With an advanced water recovery system, we can achieve better TSEC with hydrogen fuel than Jet-A.
To make hydrogen fueled engines a reality, we need to take advantage of the additional benefits of hydrogen combustion.
\subsection{Condenser Design Space Study}
\label{sub:dpqp_sweep}
% [x] TODO: AL-PA, I would emphasize that we did this sweep to explore the design space following our optimization results.
%                  Might be worth moving this into a new subsection.
% [x] TODO: AL-PA, This study wasn't motivated by a lack of literature, we want to explore the potential design space for the technology we just proposed and analyzed.
The efficiency improvement gained from water recirculation warrant a design space study for potential condenser technology.
Companies that are developing engines with water recirculation suggest that hydrogen can be used to as the heat sink for a water condenser~\cite{arpa-e_2021}.
We consider a condenser designed to fit inside the exhaust stream of a hydrogen-fueled turbofan engine.
The trade space for a condenser in the exhaust stream depends on the pressure loss and water recovery capability.
% A low pressure loss condenser would need to be developed for the water recirculation loop to be implemented.
% [x] TODO: AL-PA, dPqP is the relative total pressure loss not just a "pressure loss"

@@

% [x] TODO: AL-PA, Re-write this part to match the new figure that shows the mass flow of water extracted (per our slack convo)
The water flow rate extracted from the core stream as a function of the same pressure loss is shown in Figure~\ref{fig:dpqp_wdot}.
% Figure~\ref{fig:dpqp_wdot} shows that the optimizer finds designs that increase the available water in the core stream.
Since the water recovery fraction is always pushed to 17\%, the optimizer increases the recovered water mass flow rate to overcome the pressure loss.

@@

The hydrogen fuel would serve a double purpose in this cycle.
The cold hydrogen fuel can be used as a heat sink for the condenser and then injected into the combustor at a higher temperature, further improving efficiency~\cite{Boggia2002}.

@@

We performed a design space study to understand the technology requirements for a condenser that might operate in the proposed water recovery system.
% There are technological obstacles to overcome in order to implement this water recirculation system in a modern turbofan engine.
% The main obstacle would be how to design a condenser that can recover a portion of the exhaust water without a excessive drag pressure losses.
% Therefore, a condenser design space study was completed to show the feasible design space for such a condenser.
The results showed that a hydrogen-fueled cycle with water recovery can sustain a pressure loss of 16\% across a condenser before the efficiency benefits diminish.
% The results of the study showed that at the maximum possible water recovery fraction for hydrogen quite a large pressure loss can be imposed before the improvements are canceled out.
% Furthermore, using hydrogen as the fuel allows for designs that are not available for an engine using Jet-A.
% [x] TODO: AL-PA, I suggest you bolster the discussion of these two sentences in the results section and maybe add a citation?
The thermal properties of hydrogen provide a heat sink and thus a readily available resource for water condensation within the condenser.
The hydrogen fuel absorbs heat and is then used in the combustion process, further improving engine efficiency.
% The novel water recovery system shows promising efficiency results when used with hydrogen fuel.
Hydrogen-fueled UHBR turbofan engines can be more efficient than Jet-A counterparts when optimized considering water recovery.
Taking advantage of the beneficial properties of hydrogen is paramount for new engine technologies that can reduce emissions while improving performance.
% This water recovery system provides a potential avenue for hydrogen engine design that could help overcome the technological issues associated with hydrogen as a fuel.
